Measurement Bias from Address Aliasing

### 2014-08-14 paper.tex
Clean up description of alias counters

While optimizations such as these are good on the average case, there are corner cases. 
In particular, an event known as ``4K aliasing'' can occur when the memory addresses of a store followed by a load differ by a multiple of 4096 bytes.
A store to address 0x601020 followed by a load to address 0x821020 is an aliasing pair, because the 12-bit address suffix of 0x020 is the same in both. 
Despite being independent, in these cases the memory order subsystem generates \emph{false} dependencies, and causing the load to be reissued.
On recent architectures such as Haswell and Ivy Bridge, the number of times this happens can be counted by the following performance counter:
\begin{description}
  \item[{\small LD\_BLOCKS\_PARTIAL.ADDRESS\_ALIAS}] ``Counts the number of loads that have partial address match with preceding stores, causing the load to be reissued.'' Intel Optimization Manual~\citeyear[B.3.4.4]{OptimizationManual}
\end{description}
Older architectures such as Core2 does not have the same event available.
However, the effect is still described in the manual for these processors, and covered by the following event:
\begin{description}
  \item[{\small LOAD\_BLOCKS.OVERLAP\_STORE}] ``(...) Loads that partially overlap an earlier store, or 4-Kbyte aliased with a previous store.'' ~\citeN[Table 19-17]{Volume3B}
\end{description}

>>>>

While optimizations such as these are good on the average case, there are corner cases. 
In particular, an event known as ``4K aliasing'' can occur when the memory addresses of a store followed by a load differ by a multiple of 4096 bytes.
A store to address 0x601020 followed by a load to address 0x821020 is an aliasing pair, because the 12-bit address suffix of 0x020 is the same in both. 
Despite being independent, in these cases the memory order subsystem generates \emph{false} dependencies, and causing the load to be reissued.
The number of times this happens can be counted by the following performance counter:
\begin{describe}{{paddin}}
  \item[{\small LD\_BLOCKS\_PARTIAL.ADDRESS\_ALIAS.}] \emph{``Counts the number of loads that have partial address match with preceding stores, causing the load to be reissued.''} Intel Optimization Manual~\citeyear[B.3.4.4]{OptimizationManual}
\end{describe}
This event is listed in the manual for microarchitectures going back to Nehalem i7, including Ivy Bridge and Haswell CPUs~\cite{Volume3B}.
Older architectures such as Core 2 does not have this particular event listed, however 4K aliasing is covered by a more general event:
\begin{describe}{{paddin}}
  \item[{\small LOAD\_BLOCKS.OVERLAP\_STORE.}] \emph{``Loads that partially overlap an earlier store, or 4-Kbyte aliased with a previous store.''}~\citeN[Table 19-17]{Volume3B}
\end{describe}
